### 2018-11-21 IPCO format/FDT.tex
Abstract update

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%

\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
\usepackage{amsthm}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{Fractional Decomposition Trees: A tool for studying mixed-integer-program integrality gaps}
% \thanks{Supported by organization x.}}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Robert D. Carr\inst{1} \and
Arash Haddadan\inst{2} \and
Cynthia A. Phillips\inst{3}}

%\author{Robert Carr\inst{1}\orcidID{} \and
%Arash Haddadan\inst{2}\orcidID{} \and
%Cynthia Phillips\inst{3}\orcidID{}}

%
\authorrunning{R. Carr, A. Haddadan, C. Phillips}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
% Shorter institutions
\institute{University of New Mexico, \email{bobcarr@unm.}\\   \and
Carnegie Mellon University, \email{ahaddada@cmu.edu}\and
Sandia National Laboratories, \email{caphill@sandia.gov}}
% Longer version
%\institute{University of New Mexico, Albuquerque NM 87131, USA\\
%\email{bobcarr@unm.}\\   \and
%Carnegie Mellon University, Pittsburgh, PA 15213, USA\\
%\email{ahaddada@cmu.edu}\and
%Sandia National Laboratories, Albuquerque, NM 87185, USA\\
%\email{caphill@sandia.gov}}
%
\maketitle              % typeset the header of the con	tribution
%
\begin{abstract}
We present a new algorithm/tool for studying integrality gaps of mixed-integer linear programming formulations.
The algorithm is based on convex decomposition of scaled linear-programming relaxations. The relationship between convex decomposition and integrality gaps provides both integrality-gap information and approximate solutions. 
% The algorithm finds a feasible solution for a mixed-integer linear program.
Our algorithm runs in polynomial time and is guaranteed to find a feasible integer solution provided the
integrality gap is bounded.  Thus when the algorithm fails, it proves an unbounded integrality gap.  The algorithm also provides a lower
bound on the instance integrality gap at each step. 
% so in general it provides a suite of integer solutions.
%, this algorithm can be seen as an approximation algorithm. 
We apply our algorithm to a class of 
fractional extreme points for two traveling-salesman-like problems: 2-edge-connected spanning subgraph (2EC) and tree augmentation.
These experiments provide insight into the current gap bounds.  Furthermore, for 2EC, the approximate solutions are consistently better than the best previous
approximation algorithm due to Christofides.
%These computational experiments show that our algorithm can be used as a tool to evaluate the integrality gap of integer programs with their linear relaxation.
% Finally we provide a stronger characterization of integrality gap for a class of covering problems than that of \cite{goemans}.

\keywords{Mixed-integer linear programming  \and  Integrality gap \and convex combinations.}
\end{abstract}
%
%
%
\input{introduction.tex}
\input{binaryFDT.tex}
\input{computational}
\iffalse{
\input{strongerCV}
}\fi

%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
% \bibliographystyle{splncs04}
% \bibliography{mybibliography}
%
\begin{thebibliography}{10}

	\bibitem[BC11]{Boyd-Carr-2011}
	Sylvia Boyd and Robert Carr.
	\newblock Finding low cost tsp and 2-matching solutions using certain
	half-integer subtour vertices.
	\newblock {\em Discrete Optimization}, 8(4):525--539, 2011.

	\bibitem{carr-ravi}
	Robert Carr and R.~Ravi.
	\newblock A new bound for the 2-edge connected subgraph problem.
	\newblock In {\em Proceedings of the 6th International IPCO Conference on
		Integer Programming and Combinatorial Optimization}, pages 112--125, London,
	UK, UK, 1998. Springer-Verlag.

	\bibitem{CV}
	Robert Carr and Santosh Vempala.
	\newblock On the {H}eld-{K}arp relaxation for the asymmetric and symmetric
	traveling salesman problems.
	\newblock {\em Mathematical Programming}, 100(3):569--587, Jul 2004.

		\bibitem{fj}
	Greg N. Frederickson and Joseph Ja'Ja'.
	\newblock Approximation algorithms for several graph
	augmentation problems.
	\newblock {\em SIAM Journal on Computing}, 10(2):270--283, 1981.

		\bibitem{32gap}
	Cheriyan, Joseph and Karloff, Howard and Khandekar, Rohit and K{\"o}nemann, Jochen.
	\newblock On the integrality ratio for tree augmentation.
	\newblock {\em Operations Research Letters}, 36(4):399--401, 2008.

	\bibitem{abe}
    Anthony Alexander, Sylvia Boyd, and Paul Elliott-Magwood.
    \newblock {On the Integrality Gap of the 2-Edge Connected Subgraph Problem}.
    \newblock Technical report, University of Ottawa, Ottawa, Canada, 04 2006.

    \bibitem{TSPcompute}
    Geneviève Benoit and Sylvia Boyd.
    \newblock Finding the exact integrality gap for small traveling salesman
      problems.
    \newblock {\em Mathematics of Operations Research}, 33(4):921--931, 2008.

	\bibitem{gerard}
	G{\'e}rard Cornu{\'e}jols, Jean Fonlupt, and Denis Naddef.
	\newblock The traveling salesman problem on a graph and some related integer
	polyhedra.
	\newblock {\em Mathematical Programming}, 33(1):1--27, Sep 1985.

	\bibitem{fp1}
	Matteo Fischetti, Fred Glover, and Andrea Lodi.
	\newblock The feasibility pump.
	\newblock {\em Mathematical Programming}, 104(1):91--104, Sep 2005.

	\bibitem{fp2}
	Matteo Fischetti and Domenico Salvagnin.
	\newblock Feasibility pump 2.0.
	\newblock {\em Mathematical Programming Computation}, 1(2):201--222, Oct 2009.

	\bibitem{GJ79}
	M.R. Garey and D.S. Johnson.
	\newblock {\em Computers and Intractability -- A Guide to the Theory of
		{NP}-Completeness}.
	\newblock Freeman, San Francisco, CA, 1979.

	\bibitem{goemans}
	Michel~X. Goemans.
	\newblock Worst-case comparison of valid inequalities for the {TSP}.
	\newblock {\em Math. Program.}, 69(2):335--349, August 1995.

	\bibitem{cons-cara}
	Martin Gr{\"o}tschel, L{\'a}szl{\'o} Lov{\'a}sz, and Alexander Schrijver.
	\newblock {\em Geometric Algorithms and Combinatorial Optimization}, volume~2.
	\newblock Second corrected edition edition, 1993.

	\bibitem{jain}
	Kamal Jain.
	\newblock A factor 2 approximation algorithm for the generalized steiner
	network problem.
	\newblock In {\em Proceedings of the 39th Annual Symposium on Foundations of
		Computer Science}, FOCS '98, pages 448--, Washington, DC, USA, 1998. IEEE
	Computer Society.

	\bibitem{randomizedrounding}
	Prabhakar Raghavan and Clark~D. Tompson.
	\newblock Randomized rounding: A technique for provably good algorithms and
	algorithmic proofs.
	\newblock {\em Combinatorica}, 7(4):365--374, Dec 1987.

	\bibitem{schrijver}
	A.~Schrijver.
	\newblock {\em Combinatorial Optimization - Polyhedra and Efficiency}.
	\newblock Springer, 2003.

	\bibitem{vazirani}
	Vijay~V. Vazirani.
	\newblock {\em Approximation Algorithms}.
	\newblock Springer-Verlag, Berlin, Heidelberg, 2001.

	\bibitem{sw}
	David~P. Williamson and David~B. Shmoys.
	\newblock {\em The Design of Approximation Algorithms}.
	\newblock Cambridge University Press, New York, NY, USA, 1st edition, 2011.

	\bibitem{Wolsey1980}
	Laurence~A. Wolsey.
	\newblock {\em Heuristic analysis, linear programming and branch and bound},
	pages 121--134.
	\newblock Springer Berlin Heidelberg, Berlin, Heidelberg, 1980.

\end{thebibliography}
\newpage
\appendix

\input{2ECFDT.tex}

\end{document}