### 2019-03-22 Journal Version/computational.tex

\section{Computational experiments with FDT}\label{experiment}
We ran FDT on two covering problems: the tree augmentation problem (TAP) and 2EC. In TAP we are given a tree $T=(V,E)$, and a set of non-edges $L$ between vertices in $V$ and costs $c\in \mathbb{R}^{L}_{\geq 0}$. A feasible augmentation is $L'\subseteq L$ such that $T+L'$ is 2-edge-connected. In TAP we wish to find the minimum-cost feasible augmentation.

For $\ell\in L$, let $P_\ell$ be the set of edges in the unique path between the endpoints of $\ell$ in $T$. Let $\TAP(T,L)=\{x\in [0,1]^{L}: \sum_{\ell: e\in P_\ell}x_\ell \geq 1, \mbox{ for $e\in E$}\}$. Then $\TAP(T,L)\cap \mathbb{Z}^{L}$ is set of feasible augmentations for the instance defined by $T$ and $L$. We know $\frac{3}{2}\leq g_{\TAP}\leq 2$~\cite{fj,32gap}. We ran binary FDT on a set of 264 fractional extreme points of randomly produces instances of $\TAP$. Table \ref{tableTAP} shows FDT found solutions better than the integrality-gap lower bound for most instances.

\begin{table}[h!]
	\begin{small}
		\centering
		\begin{tabular}{c c c c c}
			\hline
			& $C\in [1.1,1.2]\;$ & $\;C\in (1.2,1.3]\;$ &
			$\;C\in (1.3,1.4]$ &\; $C\in (1.4,1.5]\;$ \\ \hline
			TAP & $36$ & $66$ & $170$ & $10$\\  \hline \\
		\end{tabular}\caption{The scale factor $C$ for FDT run on 264 randomly generated TAP instances with fractional extreme points: 138 instances have $74$ variables. The rest have $250$.} \label{tableTAP}
	\end{small}

\end{table}

We also implemented the polyhedral version of Christofides' algorithm \cite{Wolsey1980}. In particular, we implemented the $T$-join augmentation in Christofides' algorithm, in a way that minimizes the average usage of every edge across the the convex combination of spanning trees. This allows the ``real'' approximation ratio of Christofides' algorithm to be below $\frac{3}{2}$. Figure \ref{fdtvschris} shows FDT's solutions on all FEPs of order 10 are always better than those from Christofides' algorithm.

\begin{figure}[h!]
	\centering
	\includegraphics[width=6.5cm]{"fdt vs christofides".png}
	\caption{Christofides' algorithm vs FDT on all fundamental extreme points of order 10.}
	\label{fdtvschris}
\end{figure}

We ran FDT for 2EC on 963 fractional extreme points of $\EC(G)$. We enumerated all fundamental vertices of order $10$ and $12$. Table \ref{table2EC} shows that again FDT found solutions better than the integrality-gap lower bound for most instances. 
\begin{table}[h!]
\begin{small}
	\centering
	  \begin{tabular}{c c c c c}
	  	\hline
	  	& $C\in [1.08,1.11]\;$ & $\;C\in (1.11,1.14]\;$ &
               $\;C\in (1.14,1.17]$ &\; $C\in (1.17,1.2]\;$ \\ \hline
	  	2EC & $79$ & $201$ & $605$ & $78$ \\ \hline\\
	  \end{tabular}	\caption{FDT for $\EC$ implemented applied to all fundamental extreme points of order 10 or 12. A FEP of order $k$ has $\frac{3k}{2}$ variables. The lower bound on $g_{\EC}$ is $\frac{6}{5}$.}
	  \label{table2EC}
\end{small}
\end{table}