### 2020-05-15 FDT2020/FDT.tex
2020 commit

\documentclass[11pt,notitlepage]{article}

\usepackage{thumbpdf, amssymb, amsmath, amsthm, microtype, array,
graphicx, verbatim, listings, color, fancybox}
\usepackage[pdftex]{hyperref}
\usepackage{mathtools}
\usepackage{amsmath, nccmath}
\usepackage{booktabs}
\usepackage{glossaries}

\usepackage{setspace}
\usepackage{acro}
\usepackage{url}
\usepackage[linesnumbered,ruled]{algorithm2e}
\usepackage{subcaption}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{thmtools}
\usepackage{bbm}
\usepackage{pgf,tikz}
\usetikzlibrary{arrows}
\pagestyle{plain}
\usepackage[margin=1.25in]{geometry}
\usepackage{multicol,multirow}

\usepackage{mathrsfs}

\usepackage{epigraph}
\usepackage{thm-restate}
\usepackage{cleveref}

\newcommand{\ee}{e'}

\newcommand{\oc}{\mathrm{oc}}
\newcommand{\mcc}{\mathcal{C}}
\newcommand{\NP}{\mathrm{NP}}
\newcommand{\Pp}{\mathrm{P}}
\newcommand{\UGC}{\mathrm{UGC}}
\newcommand{\W}{\mathrm{W}}	
\newcommand{\FPT}{\mathrm{FPT}}
\newcommand{\opt}{\mathrm{OPT}}
\newcommand\floor[1]{\lfloor#1\rfloor}
\newcommand\ceil[1]{\lceil#1\rceil}
\newcommand{\FF}{\mathbb{F}}
\newcommand{\AF}{\mathbb{A}}
\newcommand{\BF}{\mathbb{A}}
\newcommand{\CF}{\mathbb{C}}
\newcommand{\EC}{\mathrm{2EC}}
\newcommand{\HK}{\mathrm{LP}}
\newcommand{\con}{\mathrm{Connector}}
\newcommand{\cov}{\mathrm{cov}}
\newcommand{\point}{\mathrm{end}}

\newcommand{\gap}{\alpha^{\HK}_{\EC}}
\newcommand{\Lu}{\mathcal{L}_\ell} 
\newcommand{\Rv}{\mathcal{R}_\ell}

\newcommand{\Lup}{\mathcal{L}_{\ell'}} 
\newcommand{\Rvp}{\mathcal{R}_{\ell'}}

\newcommand{\Ex}{\mathop{\mathbb{E}}}

\setlength{\tabcolsep}{0.5em} % for the horizontal padding
{\renewcommand{\arraystretch}{1.3}% for the vertical padding

\declaretheorem[name=Theorem]{thm}
\declaretheoremstyle[style=claim,qed=$\Diamond$]{claim}
\declaretheoremstyle[style=plain,qed=$\square$]{theorem}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}[thm]{Corollary}
\newtheorem{lemma}[thm]{Lemma}
\newtheorem{conjecture}{Conjecture}
\newtheorem{problem}{Open Problem}
\newtheorem{fact}[thm]{Fact}
\newtheorem{observation}[thm]{Observation}
\newtheorem{proposition}[thm]{Proposition}
\newtheorem{definition}[thm]{Definition}
\newtheorem{claim}{Claim}
\newtheorem*{notation}{Notation}
\newtheorem*{remark}{Remark}
\newtheorem{example}[thm]{Example}%[section]

%%%%%%%%%%%%%

%%%%%%%%%%%%
\DeclareMathOperator{\chord}{chord}
\DeclareMathOperator{\spp}{supp}
\DeclareMathOperator{\CC}{CC}
\DeclareMathOperator{\degree}{deg}
\DeclareMathOperator{\join}{-\;JOIN}
\DeclareMathOperator{\oddcut}{ODD-CUT}
\DeclareMathOperator{\cover}{COVER}
\DeclareMathOperator{\lp}{LP}

\DeclareMathOperator{\st}{ST}
\DeclareMathOperator{\connector}{SS}
\DeclareMathOperator{\LP}{LP}
\DeclareMathOperator{\IP}{IP}
\DeclareMathOperator{\dom}{\mathcal{D}}
\DeclareMathOperator{\conv}{conv}
\DeclareMathOperator{\subt}{{SEP}}
\DeclareMathOperator{\subtour}{{Subtour}}
\DeclareMathOperator{\hamil}{{{Hamilton}}}
\DeclareMathOperator{\tsp}{{{TSP}}}
\DeclareMathOperator{\graphtsp}{{{Graph-TSP}}}
\DeclareMathOperator{\nwtsp}{{{NW-TSP}}}
\DeclareMathOperator{\2ec}{{{2EC}}}
\DeclareMathOperator{\ecs}{{{2ECS}}}
\DeclareMathOperator{\DOMtoS}{{{DomToIP}}}
\DeclareMathOperator{\LPC}{{{LPC}}}
\DeclareMathOperator{\prun}{{{Pruning}}}

\newcommand{\vtree}{\mathrm{\textsc{-tree}}}

\newcommand{\TT}{\mathcal{T}}
\newcommand{\JJ}{\mathcal{J}}
\DeclareMathOperator{\ecss}{{{S2ECS}}}

\DeclareMathOperator{\tap}{{{TAP}}}
\DeclareMathOperator{\cut}{{{CUT}}}
\DeclareMathOperator{\permat}{{{PM}}}
\DeclareMathOperator{\nwec}{{{NW-2ECM}}}
\DeclareMathOperator{\nwecs}{{{NW-2ECS}}}
\newenvironment{cproof}
{\begin{proof}
 [Proof.]
 \vspace{-1.5\parsep}%-3.2\parsep} %%%For use with US letter
}
{\renewcommand{\qed}{\hfill $\Diamond$} \end{proof}}

\onehalfspacing
\begin{document}{\bibliographystyle{alpha}}

\begin{abstract}

\end{abstract}
\clearpage

\input{Introduction}
\input{main section}

\addcontentsline{toc}{chapter}{Bibliography}
\bibliographystyle{abbrv}
\bibliography{haddadan-ref}

\end{document}

\input{Beyond4}